An Adolescent Mini-Neptune in the Kepler Field

### 2021-08-04 paper/ms.tex
age now has uncertainties

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the {\bf $35\pm XX$\,Myr old} $\delta$~Lyr cluster.
  Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The Kepler
  photometry seems to show two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are well-poised to expand the census of
  age-dated planets, and our understanding of planetary evolution.
\end{abstract}

@@

A previous HDBScan-based clustering analysis of the Gaia data found
that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member
of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
validation of close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we opted to examine the age of the star and its putative
host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
Lyr cluster is {\bf $\approx35$\,Myr old}, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself
(Section~\ref{sec:planet}), we confirm that despite a previously
unreported binary companion, it is indeed a mini-Neptune orbiting the
roughly Solar-type primary star.  The presence of a correlation
between transit timing variations and the local light curve slope
renders any other explanation implausible.  We conclude by
highlighting some broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

\subsection{HR Diagram}
We meaured the isochrone age using an empirical approach, and also
compared against theoretical models.  The left panel of
Figure~\ref{fig:age} shows the HR diagram of stars in the \cn, the
Pleiades, IC\,2602, and the field.  The stars shown in the \cn\ are
selected based on their positions and kinematics, as discussed in
Appendix~\ref{app:kinematicselection}.  Those from the Pleiades and
IC\,2602 are adopted from \citet{cantatgaudin_gaia_2018}, and the
field stars are from the Gaia EDR3 Catalog of Nearby Stars
\citep{gaia_gcns_2021}.  We cleaned these following identical data
filtering criteria to those discussed in Appendix~B of
\citet{GaiaCollaboration2018}, except that we loosened the parallax
precision requirement to $\texttt{parallax\_over\_error}>5$.  These
filters were designed to include genuine binaries while omitting
instrumental artifacts.  We then corrected for extinction by querying
the 3-D maps of \citet{capitanio_threedimensional_2017} and
\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
and by then using the extinction coefficients $k_X\equiv A_X/A_0$
computed by \citet{GaiaCollaboration2018}, while assuming that $A_0 =
3.1 E(B-V)$.  For IC\,2602, the Pleaides, and the \cn, this procedure
yielded a respective mean and standard deviation for the reddening of
$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.  These
values agree reasonably well with previously reported values from the
literature \citep[{\it
e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019},
and have the benefit of being sensitive to effects such as
differential reddening that may be present in the dust maps.

@@

The immediate impression from the left panel of Figure~\ref{fig:age}
is that the \cn\ and IC\,2602 share approximately the same isochronal
age, since they overlap.  In our exploration, we also compared against
$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and subsets
of the Sco OB2 association ($\approx10$-$15$\,Myr;
\citealt{damiani_stellar_2019}), and confirmed that the
pre-main-sequence M dwarfs were intermediate between these two
clusters.

% FIXME FIXME
We quantified this impression following the empirical fitting procedure
discussed by \citet{gagne_mutau_2020}.

>>>>

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.
  Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The Kepler
  photometry seems to show two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are well-poised to expand the census of
  age-dated planets, and our understanding of planetary evolution.
\end{abstract}

@@

A previous HDBScan-based clustering analysis of the Gaia data found
that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member
of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
validation of close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we opted to examine the age of the star and its putative
host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
Lyr cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself
(Section~\ref{sec:planet}), we confirm that despite a previously
unreported binary companion, it is indeed a mini-Neptune orbiting the
roughly Solar-type primary star.  The presence of a correlation
between transit timing variations and the local light curve slope
renders any other explanation implausible.  We conclude by
highlighting some broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

\subsection{Color-Magnitude Diagram}
We meaured the isochrone age using an empirical approach, and also
compared against theoretical models.  The left panel of
Figure~\ref{fig:age} shows the color-magnitude diagram (CMD) of stars in the \cn, the
Pleiades, IC\,2602, and the field.  The stars shown in the \cn\ are
selected based on their positions and kinematics, as discussed in
Appendix~\ref{app:kinematicselection}.  Those from the Pleiades and
IC\,2602 are adopted from \citet{cantatgaudin_gaia_2018}, and the
field stars are from the Gaia EDR3 Catalog of Nearby Stars
\citep{gaia_gcns_2021}.  We cleaned these following identical data
filtering criteria to those discussed in Appendix~B of
\citet{GaiaCollaboration2018}, except that we loosened the parallax
precision requirement to $\varpi/\sigma_\varpi>5$.  These
filters were designed to include genuine binaries while omitting
instrumental artifacts.  We then corrected for extinction by querying
the 3-D maps of \citet{capitanio_threedimensional_2017} and
\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
and by then using the extinction coefficients $k_X\equiv A_X/A_0$
computed by \citet{GaiaCollaboration2018}, while assuming that $A_0 =
3.1 E(B-V)$.  For IC\,2602, the Pleaides, and the \cn, this procedure
yielded a respective mean and standard deviation for the reddening of
$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.  These
values agree reasonably well with previously reported values from the
literature \citep[{\it
e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019}.

@@

The immediate impression from the left panel of Figure~\ref{fig:age}
is that the \cn\ and IC\,2602 share approximately the same isochronal
age, since they overlap.  In our exploration, we also compared against
$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and the
Upper-Centaurus-Lupus component of the Sco OB2 association (UCL;
$\approx10$-$15$\,Myr; \citealt{damiani_stellar_2019}).  We confirmed
that the pre-main-sequence M dwarfs of the \cn\ were intermediate
between these two clusters.  To quantify this into an age measurement
with an uncertainty, we adopted the empirical procedure
advanced by \citet{gagne_mutau_2020} in their Section~6.3.  In brief,
we parametrized the pre-main-sequence locus of each of a set of
reference clusters using splines, and the locus of the target \cn\ was
then fitted as a piecewise linear combination of these reference
clusters.  As reference clusters, we adopted members of UCL, IC\,2602, and the
Pleaides from \citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}.
We removed binaries from each set by requiring
$\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$ below the
80$^{\rm th}$ percentile of each cluster's distribution, and by
manually excluding stars that were obvious photometric binaries.  We
then passed a moving box average and standard deviation across the
CMD in 0.10\,mag bins, fitted a univariate spline to the binned
values, and assembled a piecewise grid of hybrid isochrones spanning
the ages between UCL to the Pleiades following Equations~6 and~7 from
\citet{gagne_mutau_2020}.

The ages calculated from this procedure depend on the ages assumed for
each reference cluster.
In our case, the IC\,2602 cluster is most important.
As noted by \citet{bouma_cluster_2020}, plausible ages for this
cluster span 30 to 46\,Myr, with slightly older ages being
preferred by the lithium-depletion-boundary (LDB) measurements and younger
ages by the main-sequence turn-off
\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
If we were to adopt the lower 30\,Myr age for IC\,2602, the \cn\ would be
$31^{+5}_{-4}$\,Myr old.
For the converse case of the 46\,Myr boundary, the \cn\ would be
$44^{+8}_{-7}$\,Myr old.
Taking an intermediate 35\,Myr age as our reference, 
the resulting age for the \cn\ is
$35^{+6}_{-5}$\, Myr, though it should be understood based on the
systematic uncertainties that an age of up to $\approx50$\,Myr remains
plausible.  A detailed analysis of the LDB and main-sequence
turn-off in the \cn\ could help clarify the issue, but both are beyond
our scope.
